Detection and Classification of Supernova Gravitational Waves Signals: A Deep Learning Approach

### 2020-03-25 main.tex
comments incoporated version
revised version of the paper

\begin{abstract}
We demonstrate the application of a convolutional neural network to the
gravitational wave signals from core collapse supernovae. Using simulated time
series of gravitational wave detectors, we show that based on the explosion
mechanisms, a convolutional neural network can be used to detect and classify
the gravitational wave signals buried in noise. For the waveforms used in the
training of the convolutional neural network, our results suggest that a
network of advanced LIGO, advanced VIRGO and KAGRA, or a network of LIGO A+,
advanced VIRGO and KAGRA is likely to detect a magnetorotational core collapse
supernovae within the Large and Small Magellanic Clouds, or a Galactic event if
the explosion mechanism is the neutrino-driven mechanism.  By testing the
convolutional neural network with waveforms not used for training, we show that
the true alarm probabilities are $52\%$ and $83\%$ at $60$ kpc for waveforms
$\text{R3E1AC}$ and $\text{R4E1FC\_L}$.  For waveforms $\text{s}20$ and
$\text{SFHx}$ at $10$ kpc, the true alarm probabilities are $70\%$ and $93\%$
respectively. All at false alarm probability equal to $10\%$. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%
% general background on GWs
%
Since 2015 when LIGO made the first direct observation of \acp{GW} from the
merger of a binary black hole~\cite{abbott2016observation}, there have been
numerous observations of \acp{GW} from similar systems in the subsequent
observation runs of LIGO and VIRGO~\cite{abbott2016gw151226,
abbott2017gw170608, abbott2017gw170814}. These discoveries represent a crucial
milestone in \ac{GW} astronomy and have opened up a new window on the universe.
More recently, LIGO and VIRGO have observed the \acp{GW} from a binary neutron
star merger~\cite{abbott2017gw170817, abbott2017gravitational, abbott2017multi}
where the \acp{GW} and an associated gamma-ray burst were observed
simultaneously. Other counterparts across the electromagnetic spectrum were
also observed by later follow-up observations~\cite{abbott2017multi}. In the
near future, many more observations of \acp{GW} from similar \ac{CBC} systems
can be expected as KAGRA starts joint observations with LIGO and
VIRGO~\cite{aso2013interferometer, somiya2012detector, abbott2018prospects}.

%
% intro to core-collapse supernovae
%
In addition to \acp{CBC}, massive stars with $10-100 \text{M}_\odot $ at
zeros-age main sequence ending their lives by becoming 
%with core of $10-100 \text{M}_\odot $ prior to collapse 
\acp{CCSN} are also considered to be potential sources to the second generation
detectors such as \ac{aLIGO}~\cite{aasi2015advanced},
\ac{AdVirgo}~\cite{acernese2014advanced} and KAGRA
interferometers~\cite{aso2013interferometer, gossan2016observing,
abbott2016first, abbott2019optically}. It is currently not entirely clear to astronomers how such
massive stars become supernovae. The basic theory of the explosion, confirmed
by the neutrino events observed from SN1987A~\cite{sato1987analysis}, begins
with a massive star at the final stage of its life forming a core that is
composed of iron nuclei after it has burned all its stellar fuel via fusion
reactions. The iron core is supported by the pressure of relativistic
degenerate electrons and if the mass of the core exceeds the effective
Chandrasekhar mass~\cite{baron1990effect, bethe1990supernova}, core collapse
will ensue and continue until the core reaches nuclear density. The nuclear
equation of state will then stiffens by the strong nuclear force above the
nuclear density and stops the core collapse. The inner core will bounce back
and a shock wave will be sent through the infalling matter.  By losing energy
to the dissociation of the iron nuclei and to neutrino cooling, the shock wave
will stall. For the star to become a supernova, the shock wave will need to be
revived~\cite{o2011black}. However the mechanism via which this occurs and the
supernova explosion is caused has been the subject of intense study and remains
an unsolved problem.

@@

outputs of the \ac{CNN} (the probabilities of the inferred classes in our case) are used as an input to the loss-function that is associated with the entire network.  

@@

distinguishing supernova signals in backgound \ac{GW} detector noise vs detectr

@@

%
% describe the neutrino waveforms
%
For the neutrino-driven mechanism, we employ the waveforms
from~\cite{10.1093mnrasstz990, kuroda2017correlated, muller2012parametrized,
powell2019gravitational, radice2019characterizing, yakunin2015gravitational,
yakunin2017gravitational, ott2009gravitational, murphy2009model,
ott2013general}\footnote{\href{https://wwwmpa.mpa-garching.mpg.de/ccsnarchive/data/Andresen2019/}
{https://wwwmpa.mpa-garching.mpg.de/ccsnarchive/
data/Andresen2019/ }}$^{,}$\footnote{\href{ https://www.astro.princeton.edu/~burrows/gw.3d/  }
{ https://www.astro.princeton.edu/~burrows/gw.3d/  }}. The simulations in~\cite{10.1093mnrasstz990,
kuroda2017correlated, muller2012parametrized, powell2019gravitational,
radice2019characterizing, yakunin2017gravitational, ott2013general} were from 3D modelling of
the supernovae while the simulations in~\cite{yakunin2015gravitational,
ott2009gravitational, murphy2009model} were 2D. These
simulations cover a wide range of progenitor masses from $9\text{M}_\odot$ to
$60\text{M}_\odot$ and the specific progenitor masses of the simulations are
shown in Table~\ref{table:waveforms}. Examples of the waveforms for
both mechanisms are shown in Figs.~\ref{fig:magwaveforms} and
\ref{fig:neuwaveforms}.

@@

%
% plot of example magneto waveforms
%
\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:magsample1}
            \includegraphics[width=0.315\textwidth]{abd_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:magsample2}
            \includegraphics[width=0.315\textwidth]{Dim_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:magsample3}
            \includegraphics[width=0.315\textwidth]{richer_wave.png}%
        }
    \end{center}
    \caption{Examples of simulated waveforms for which the explosion is modelled by
the magnetorotational mechanism. The left panel shows a waveform
from~\cite{abdikamalov2014measuring}. The progenitor is $12\text{M}_\odot$ with
differential rotation parameter $A=10^4~\text{km}$. The initial angular velocity
at the core $\Omega_c$ is $5.0~\text{rads/s}$. The middle panel shows a waveform
from~\cite{dimmelmeier2008gravitational}. The simulation corresponds to a
progenitor of $15\text{M}_\odot$ with $A=10^5$ km and $\Omega_c =
4.56~\text{rads/s}$.  The right panel shows a waveform
from~\cite{richers2017equation} where the progenitor is $12\text{M}_\odot$, $A$
and $\Omega_c$ are $10^4$ km and $3.0~\text{rad/s}$ respectively. In all panels,
only the $h_{+}$ polarisations are shown because the simulations are
axis-symmetric or 2D, and therefore described by only one polarisation.  The
sources are assumed to be at a distance of $10$ kpc. The x-axes show the time after 
core bounce.\label{fig:magwaveforms}} 
\end{figure*}
%
% plot of example neutrino waveforms
%
\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:neusample1}
            \includegraphics[width=0.315\textwidth]{and_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:neusample2}
            \includegraphics[width=0.315\textwidth]{Radice_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:neusample3}
            \includegraphics[width=0.315\textwidth]{Yak_wave.png}%
        }
    \end{center}
    \caption{Examples of simulated waveforms for which the explosion mechanism is
modelled by the neutrino-driven mechanism. From the left to the right, the
panels show waveforms from~\cite{10.1093mnrasstz990, 
radice2019characterizing, yakunin2017gravitational}, with progenitors of masses equal to
$15\text{M}_\odot$, $15\text{M}_\odot$ and $60\text{M}_\odot$ respectively.
The sources are assumed to be at $10$ kpc from earth. The x-axes show the time after core bounce.
\label{fig:neuwaveforms}} 
\end{figure*}
%
% discuss the reason of using multiple waveform families
%
Introducing waveforms from a variety of studies introduces a distribution of
the waveforms that covers a larger parameter space, making the data samples
harder for the \ac{CNN} to learn. As a result, the performance of the trained
\ac{CNN} on the testing samples may appear to be worse than that if the data
were generated using only a subset of the waveforms. However, a larger
parameter space provides the advantage of forcing the \ac{CNN} to learn the
features common among the waveforms rather than simply a subset of the
waveforms.  The trained network can thus perform better when presented with
waveforms with unexpected features, which is more likely in reality.

@@

%
% data augmentation
%
As a general rule, the performance of a \ac{CNN} can always be enhanced by
increasing the volume of training data. Therefore we augment our data set by
iterating over $\bold{S}(t)$ until for each waveform in $\bold{S}(t)$, we have
$31$ data samples $\bold{d}$ generated using the procedure described above.
This step is essential for the \ac{CNN} to learn the features of the waveforms
and to identify them effectively under different noise scenarios and starting
times.  After that, we generate independent noise realisations for the
background noise class. The entire data set $\bold{D}$ has
roughly $l = 1.8 \times 10^{5}$ data samples, where each of the 3 classes
contains approximately $6\times10^{4}$ data samples.
%
%\begin{equation}\label{eq:finaldata}
%\bold{D}(\boldsymbol{\alpha}, \boldsymbol{\delta}, \boldsymbol{\delta_t}, \bold{t}, \bold{x}) =  \left(\begin{array}{c}
%\bold{d}_1(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_o(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_l(t) \\
%\end{array}\right).
%\end{equation}
%
%In the above equation, the subscript $o$ is the sum of the data samples in
%which there is a \ac{GW} signal in the time series.~\cm{this notation is very
%confusing and not really required. Why do you have to stack the waveforms up
%like this? Having to define things like $o$ and $l$ seems messy.} The final
%step is to whiten $\bold{D}$ using the power spectral densities. 
%
% the general training procedure
%
In Fig.~\ref{fig:sample}, we show a representative example of a pre-processed
input time-series for a single detector. A data sample consists of such a
time-series containing a signal from the same source or simply background noise
from all the detectors in the network. The data samples are then split into 3
groups, with $1\times10^{4}$ samples being randomly selected for validation,
$10\%$ of the remainder for testing and $90\%$ for training. When the training
is finished for a dataset generated at a given distance, the above described
procedure will be repeated for another distance with a different \ac{CNN} of 
the same architecture until the training for all
distances have been carried out for a \ac{GW} detector network. The entire
procedure is then repeated for another \ac{GW} detector network. 
As mentioned in the introduction, the \ac{CNN} will be trained for the two networks of
\ac{GW} detectors presented in Table~\ref{table:network}. For the remaining of
the paper, we will use their acronyms to refer to the networks, i.e., HLVK and H+L+VK.

@@

%
% discuss the accuracy as a function of distance (Fig 6)
%
We also show the classification efficiency of the \ac{CNN}  as a function of
distance.  This is done by fixing the \ac{FAP} and plotting the \ac{TAP}.  The
results for three chosen \acp{FAP} values are shown in Fig.~\ref{fig:eff}.  In
this figure, a similar trend is seen that magnetorotational signals are easier
for the \ac{CNN} to identify than neutrino-driven signals at a given distance.
For magnetorotational signals from sources located at $50$ kpc and a \ac{FAP}
of $10\%$, the \ac{CNN} achieves a \ac{TAP} of $94\%$ and $82\%$ for the
networks H+L+VK and HLVK respectively. At more restrict \acp{FAP} such as
$0.1\%$, the \ac{CNN} still achieves \acp{TAP} $86\%$ and $68\%$ for sources at
the same distance for the two networks respectively.  For sources that are
located at a slightly further distance of $60$ kpc, both networks achieve a
\ac{TAP} of close to or larger than $80\%$ at a \ac{FAP} of $10\%$. Such a range
includes distances associated with the Large and Small Magellanic Clouds and
covers the satellite galaxies in between~\cite{karachentsev2004catalog,
belokurov2007cats}.  For sources at $100$ kpc, the \acp{TAP} are close to or
larger than $60\%$ for H+L+VK for all chosen \acp{FAP} values, and is $73\%$ if
the \ac{FAP} is $10\%$. Even for sources at $150$ and $200$ kpc, the \acp{TAP}
are $54\%$ and $38\%$ respectively for the same \ac{FAP}, indicating that with
such a \ac{GW} network, it is possible to detect magnetorotational \ac{CCSN}
signals out to such a distance. On the other hand, it is more difficult for the
\ac{CNN} to detect and classify the neutrino-driven signals, due to their
weaker amplitudes. Nonetheless, for sources at $10$ kpc, the \ac{CNN} achieves
a \ac{TAP} of $76\%$ and $55\%$ for H+L+VK and HLVK respectively if the
\ac{FAP} is $10\%$.  This means that for a \ac{GW} from a Galactic \ac{CCSN}
signal it is possible to detect and classify it with either of these \ac{GW}
detector networks.

%
% the TAP as a function of distance
%
\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:eff1}
            \includegraphics[width=0.475\textwidth]{Neutrino_efficiency.png}
        }\quad
        \subfigure[]{
            \label{fig:eff2}
            \includegraphics[width=0.475\textwidth]{Magnetarotational_efficiency.png}
        }
    \end{center}
    \caption{Efficiency curves showing the classification ability of the
\ac{CNN} as a function of distance for both mechanisms and networks.  The left
panel shows the results for the neutrino-driven mechanism, while the right
shows those for the magnetorotational mechanism. In both panels, the solid
lines show the results for the network H+L+VK, and the dashed lines for HLVK.
Three \acp{FAP} values are chosen, blue for \ac{FAP} $=10\%$, green for \ac{FAP}
$=1\%$, red for \ac{FAP} $=0.1\%$.\label{fig:eff}}
\end{figure*}

@@

In practice, since the output of a \ac{CNN} is probabilities indicating how
likely the input belongs to each of the classes, it may be desired to set a
threshold probability on which the decision whether the input belong to a class is made.
For example, an input will be assigned into a certain class if the
corresponding probability is larger than the pre-selected threshold probability. In such a
scenario, the \ac{FAP} and \ac{TAP} would be affected by the choice of the
threshold. We show such a result in Fig.~\ref{fig:ROCfixed}, where we employed
a threshold probability of $50\%$. This value is chosen because for lower values, 
there could be more than a class with predicted probabilities from the \ac{CNN} 
larger than the threshold, while higher values could potentially rule out inputs 
that may otherwise be correctly identified.
For H+L+VK and magnetorotational signals at $10$ kpc, the \ac{TAP} is $99\%$. If the distance
is extended to $80$ kpc, the \ac{TAP} is still close to $80\%$.  For the
largest distances tested in this work, $150$ and $200$ kpc, the \acp{TAP} are
$43\%$ and $29\%$ respectively. 
For HLVK, the \ac{TAP} is $98\%$ and $49\%$ for sources at $10$ and
$80$ kpc respectively.
%The \acp{TAP} decrease to $10\%$ and $9\%$ for sources at $150$ and $200$ kpc.
For all distances, the \ac{CNN} maintains a \ac{FAP} no larger than $4\%$ for
both networks regarding magnetorotational waveforms. For neutrino-driven
signals, we show that H+L+VK has a \acp{TAP} of $66\%$ at $10$ kpc while it
is $55\%$ at $10$ kpc for HLVK.  Both of the networks have \acp{FAP} close to
or less than $20\%$. It should be noted that the results presented in this section are
averaged over all the data samples in the testing samples. The performance on
any individual waveform may vary depending on the morphology and the amplitudes
of the waveform.

>>>>

\begin{abstract}
We demonstrate the application of a convolutional neural network to the
gravitational wave signals from core collapse supernovae. Using simulated time
series of gravitational wave detectors, we show that based on the explosion
mechanisms, a convolutional neural network can be used to detect and classify
the gravitational wave signals buried in noise. For the waveforms used in the
training of the convolutional neural network, our results suggest that a
network of advanced LIGO, advanced VIRGO and KAGRA, or a network of LIGO A+,
advanced VIRGO and KAGRA is likely to detect a magnetorotational core collapse
supernovae out to the Large or even Small Magellanic Clouds. A Galactic event 
is likely to be detectable with a network of H+L+VK if the explosion mechanism 
is the neutrino-driven mechanism.  By testing the convolutional neural network 
with waveforms not used for training, we show that the true alarm probabilities 
are $50\%$ and $73\%$ at $50$ kpc for waveforms $\text{R3E1AC}$ and 
$\text{R4E1FC\_L}$ for networks of HLVK and H+L+VK respectively.  
For waveforms $\text{s}20$ and $\text{SFHx}$ at $10$ kpc, 
the true alarm probabilities are $64\%$ for a network of H+L+VK. 
All at false alarm probability equal to $1\%$. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%
% general background on GWs
%
In 2015, LIGO made the first direct observation of \acp{GW} from the
merger of a binary black hole~\cite{abbott2016observation} 
with the detector of \ac{aLIGO}~\cite{aasi2015advanced}.
Since then and after VIRGO joined the observations 
with its \ac{AdVirgo} detector~\cite{abbott2016gw151226,
abbott2017gw170608, abbott2017gw170814}, there have been
numerous observations of \acp{GW} from similar systems in the subsequent
observation runs of LIGO and VIRGO. These discoveries represent a crucial
milestone in \ac{GW} astronomy and have opened up a new window on the universe.
More recently, LIGO and VIRGO have observed the \acp{GW} from a binary neutron
star merger~\cite{abbott2017gw170817, abbott2017gravitational, abbott2017multi}
where the \acp{GW} and an associated gamma-ray burst were observed
simultaneously. Other counterparts across the electromagnetic spectrum were
also observed by later follow-up observations~\cite{abbott2017multi}. In the
near future, many more observations of \acp{GW} from similar \ac{CBC} systems
can be expected as the KAGRA interferometer starts joint observations with LIGO and
VIRGO~\cite{aso2013interferometer, somiya2012detector, abbott2018prospects, gossan2016observing,
abbott2016first, abbott2019optically}.

%
% intro to core-collapse supernovae
%
In addition to \acp{CBC}, massive stars with $10-100 \text{M}_\odot $ at
zeros-age main sequence ending their lives by becoming 
%with core of $10-100 \text{M}_\odot $ prior to collapse 
\acp{CCSN} are also considered to be potential sources to
\ac{aLIGO}, \ac{AdVirgo} and KAGRA interferometers. 
It is currently not entirely clear to astronomers how such
massive stars become supernovae. The basic theory of the explosion, confirmed
by the neutrino events observed from SN1987A~\cite{sato1987analysis}, begins
with a massive star at the final stage of its life forming a core that is
composed of iron nuclei after it has burned all its stellar fuel via fusion
reactions. The iron core is supported by the pressure of relativistic
degenerate electrons and if the mass of the core exceeds the effective
Chandrasekhar mass~\cite{baron1990effect, bethe1990supernova}, core collapse
will ensue and continue until the core reaches nuclear density. The nuclear
equation of state will then stiffens by the strong nuclear force as the core exceeds
nuclear density and stops the core collapse. The inner core will bounce back
and a shock wave will be sent through the infalling matter.  By losing energy
to the dissociation of the iron nuclei and to neutrino cooling, the shock wave
will stall. For the star to become a supernova, the shock wave will need to be
revived~\cite{o2011black}. However the mechanism via which this occurs and the
supernova explosion is caused has been the subject of intense study and remains
an unsolved problem.

@@

outputs of the \ac{CNN} (the probabilities of the inferred classes in our case) 
are used as an input to the loss-function that is associated with the entire network.  

@@

distinguishing supernova signals in background \ac{GW} detector noise vs detector

@@

%
% describe the neutrino waveforms
%
For the neutrino-driven mechanism, we employ the waveforms
from~\cite{10.1093mnrasstz990, kuroda2017correlated, muller2012parametrized,
powell2019gravitational, radice2019characterizing, yakunin2015gravitational,
yakunin2017gravitational, ott2009gravitational, murphy2009model,
ott2013general}\footnote{\href{https://wwwmpa.mpa-garching.mpg.de/ccsnarchive/data/Andresen2019/}
{https://wwwmpa.mpa-garching.mpg.de/ccsnarchive/
data/Andresen2019/ }}$^{,}$\footnote{\href{ https://www.astro.princeton.edu/~burrows/gw.3d/  }
{ https://www.astro.princeton.edu/~burrows/gw.3d/  }}. The simulations in~\cite{10.1093mnrasstz990,
kuroda2017correlated, muller2012parametrized, powell2019gravitational,
radice2019characterizing, yakunin2017gravitational, ott2013general} were from 3D modelling of
the supernovae. 
In particular, the waveforms from~\cite{10.1093mnrasstz990, muller2012parametrized,  ott2013general} assuming an observer both along the equator and at the pole
are available, 
the waveforms from~\cite{kuroda2017correlated, powell2019gravitational} are available for an observer at the pole, and the waveforms from~\cite{yakunin2017gravitational,
radice2019characterizing} for an equatorial observer.
The simulations in~\cite{yakunin2015gravitational,
ott2009gravitational, murphy2009model} were 2D.   These
simulations cover a wide range of progenitor masses from $9\text{M}_\odot$ to
$60\text{M}_\odot$ and the specific progenitor masses of the simulations are
shown in Table~\ref{table:waveforms}. Examples of the waveforms for
both mechanisms are shown in Figs.~\ref{fig:magwaveforms} and
\ref{fig:neuwaveforms}.

@@

%
% plot of example magneto waveforms
%
\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:magsample1}
            \includegraphics[width=0.315\textwidth]{abd_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:magsample2}
            \includegraphics[width=0.315\textwidth]{Dim_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:magsample3}
            \includegraphics[width=0.315\textwidth]{richer_wave.png}%
        }
    \end{center}
    \caption{Examples of simulated waveforms for which the explosion is modelled by
the magnetorotational mechanism. The left panel shows a waveform 
from~\cite{abdikamalov2014measuring} for an equatorial observer. The progenitor is $12\text{M}_\odot$ with
differential rotation parameter $A=10^4~\text{km}$. The initial angular velocity
at the core $\Omega_c$ is $5.0~\text{rads/s}$. The middle panel shows a waveform
from~\cite{dimmelmeier2008gravitational}. The simulation corresponds to a
progenitor of $15\text{M}_\odot$ with $A=10^5$ km and $\Omega_c =
4.56~\text{rads/s}$.  The right panel shows a waveform
from~\cite{richers2017equation} where the progenitor is $12\text{M}_\odot$, $A$
and $\Omega_c$ are $10^4$ km and $3.0~\text{rad/s}$ respectively. In all panels,
only the $h_{+}$ polarisations are shown because the simulations are
axis-symmetric or 2D, and therefore described by only one polarisation.  The
sources are assumed to be at a distance of $10$ kpc. The x-axes show the time after 
core bounce. The waveforms shown in all panels are for an observer in the equatorial plane.\label{fig:magwaveforms}} 
\end{figure*}
%
% plot of example neutrino waveforms
%
\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:neusample1}
            \includegraphics[width=0.315\textwidth]{and_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:neusample2}
            \includegraphics[width=0.315\textwidth]{Radice_wave.png}%
        }\quad
        \subfigure[]{
            \label{fig:neusample3}
            \includegraphics[width=0.315\textwidth]{Yak_wave.png}%
        }
    \end{center}
    \caption{Examples of simulated waveforms for which the explosion mechanism is
modelled by the neutrino-driven mechanism. From the left to the right, the
panels show waveforms from~\cite{10.1093mnrasstz990, 
radice2019characterizing, yakunin2017gravitational}, with progenitors of masses equal to
$15\text{M}_\odot$, $15\text{M}_\odot$ and $60\text{M}_\odot$ respectively.
For the left and middle panels, the observer is assumed to be along the pole, 
while an observer along the equator is assumed for the right panel.
The sources are assumed to be at $10$ kpc from earth. The x-axes show the time after core bounce.
The waveform shown in the left panel is for a polar observer, while the other panels show that 
for an equatorial observer.
\label{fig:neuwaveforms}} 
\end{figure*}
%
% discuss the reason of using multiple waveform families
%
Introducing waveforms from a variety of studies introduces a distribution of
the waveforms that covers a larger parameter space, making the data samples
harder for the \ac{CNN} to learn. As a result, the performance of the trained
\ac{CNN} on the testing samples may appear to be worse than that if the data
were generated using only a subset of the waveforms. However, a larger
parameter space provides the advantage of forcing the \ac{CNN} to learn the
features common among the waveforms rather than simply a subset of the
waveforms.  The trained network can thus perform better when presented with
waveforms with unexpected features, which is more likely in reality.

@@

%
% data augmentation
%
As a general rule, the performance of a \ac{CNN} can always be enhanced by
increasing the volume of training data. Therefore we augment our data set by
iterating over $\bold{S}(t)$ until for each waveform in $\bold{S}(t)$, we have
$31$ data samples $\bold{d}$ generated using the procedure described above.
This step is essential for the \ac{CNN} to learn the features of the waveforms
and to identify them effectively under different noise scenarios and starting
times.  After that, we generate independent noise realisations for the
background noise class. The entire data set $\bold{D}$ has
roughly $l = 3.0 \times 10^{5}$ data samples, where each of the 3 classes
contains approximately $1.0\times10^{5}$ data samples.
%
%\begin{equation}\label{eq:finaldata}
%\bold{D}(\boldsymbol{\alpha}, \boldsymbol{\delta}, \boldsymbol{\delta_t}, \bold{t}, \bold{x}) =  \left(\begin{array}{c}
%\bold{d}_1(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_o(\alpha, \delta, \delta_t, t, x) \\
%\vdots \\
%\bold{d}_l(t) \\
%\end{array}\right).
%\end{equation}
%
%In the above equation, the subscript $o$ is the sum of the data samples in
%which there is a \ac{GW} signal in the time series.~\cm{this notation is very
%confusing and not really required. Why do you have to stack the waveforms up
%like this? Having to define things like $o$ and $l$ seems messy.} The final
%step is to whiten $\bold{D}$ using the power spectral densities. 
%
% the general training procedure
%
In Fig.~\ref{fig:sample}, we show a representative example of a pre-processed
input time-series for a single detector. A data sample consists of such a
time-series containing a signal from the same source or simply background noise
from all the detectors in the network. The data samples are then split into 3
groups, with $1.0\times10^{4}$ samples being randomly selected for validation.
Of the remainder, $\sim 1.7\times10^{5}$ for training and $\sim 1.2\times10^{5}$ for tesing. 
The large number of testing samples is chosen so that low values of \ac{FAP} can be obtained 
(see section \ref{sec:result} and section \ref{sec:unseen}).
When the training is finished for a dataset generated at a given distance, the above described
procedure will be repeated for another distance with a different \ac{CNN} of 
the same architecture until the training for all
distances have been carried out for a \ac{GW} detector network. The entire
procedure is then repeated for another \ac{GW} detector network. 
As mentioned in the introduction, the \ac{CNN} will be trained for the two networks of
\ac{GW} detectors presented in Table~\ref{table:network}. For the remaining of
the paper, we will use their acronyms to refer to the networks, i.e., HLVK and H+L+VK.

@@

%
% discuss the accuracy as a function of distance (Fig 6)
%
We also show the classification efficiency of the \ac{CNN}  as a function of
distance.  This is done by fixing the \ac{FAP} and plotting the \ac{TAP}.  The
results for three chosen \acp{FAP} values are shown in Fig.~\ref{fig:eff}.  In
this figure, a similar trend is seen that magnetorotational signals are easier
for the \ac{CNN} to identify than neutrino-driven signals at a given distance.
For magnetorotational signals from sources located at $50$ kpc and a \ac{FAP}
of $1\%$, the \ac{CNN} achieves a \ac{TAP} of $87\%$ and $72\%$ for the
networks H+L+VK and HLVK respectively. At more restrict \acp{FAP} such as
$0.01\%$, the \ac{CNN} still achieves \acp{TAP} of $75\%$ and $62\%$ for sources at
the same distance for the two networks respectively.  For sources that are
located at a slightly further distance of $60$ kpc, the networks achieve a
\ac{TAP} of $82\%$ (H+L+VK) or $60\%$ (HLVK) at a \ac{FAP} of $1\%$. Such a range
includes distances associated with the Large and Small Magellanic Clouds and
covers the satellite galaxies in between~\cite{karachentsev2004catalog,
belokurov2007cats}.  For sources at $100$ kpc, the \acp{TAP} are no less than $50\%$ 
for H+L+VK for all chosen \acp{FAP} values, and is $61\%$ if
the \ac{FAP} is $1\%$. These numbers indicate that with
such a \ac{GW} network, it is possible to detect magnetorotational \ac{CCSN}
signals out to such a distance. For sources at $150$ and $200$ kpc, the \acp{TAP}
drop to $39\%$ and $21\%$ respectively for \ac{FAP} equal to $1\%$.
On the other hand, it is more difficult for the
\ac{CNN} to detect and classify the neutrino-driven signals, due to their
weaker amplitudes. For example, for sources at $10$ kpc, the \ac{CNN} achieves
a \ac{TAP} of $53\%$ and $22\%$ for H+L+VK and HLVK respectively if the
\ac{FAP} is $1\%$.  This means that for a \ac{GW} from a Galactic \ac{CCSN}
signal it is possible to detect and classify it with a network of H+L+VK.

%
% the TAP as a function of distance
%
\begin{figure*}
     \begin{center}
        \subfigure[]{
            \label{fig:eff1}
            \includegraphics[width=0.475\textwidth]{Neutrino_efficiency.png}
        }\quad
        \subfigure[]{
            \label{fig:eff2}
            \includegraphics[width=0.475\textwidth]{Magnetarotational_efficiency.png}
        }
    \end{center}
    \caption{Efficiency curves showing the classification ability of the
\ac{CNN} as a function of distance for both mechanisms and networks.  The left
panel shows the results for the neutrino-driven mechanism, while the right
shows those for the magnetorotational mechanism. In both panels, the solid
lines show the results for the network H+L+VK, and the dashed lines for HLVK.
Three \acp{FAP} values are chosen, blue for \ac{FAP} $=1\%$, green for \ac{FAP}
$=0.1\%$, red for \ac{FAP} $=0.01\%$.\label{fig:eff}}
\end{figure*}

@@

In practice, since the output of a \ac{CNN} is probabilities indicating how
likely the input belongs to each of the classes, it may be desired to set a
threshold probability on which the decision whether the input belong to a class is made.
For example, an input will be assigned into a certain class if the
corresponding probability is larger than the pre-selected threshold probability. In such a
scenario, the \ac{FAP} and \ac{TAP} would be affected by the choice of the
threshold. We show such a result in Fig.~\ref{fig:ROCfixed}, where we employed
a threshold probability of $95\%$. 
This value is chosen so that low values of \ac{FAP} can be maintained throughout all tested distances
as illustrated by the right y-axes of the figures.
%because for lower values, 
%there could be more than a class with predicted probabilities from the \ac{CNN} 
%larger than the threshold, while higher values could potentially rule out inputs 
%that may otherwise be correctly identified.
For H+L+VK and magnetorotational signals at $10$ kpc, the \ac{TAP} is $97\%$. If the distance
is extended to $100$ kpc, the \ac{TAP} is still larger than $50\%$. 
For HLVK, the \ac{TAP} is $95\%$ for sources at $10$ kpc and drops to $54\%$ at
$60$ kpc.
%The \acp{TAP} decrease to $10\%$ and $9\%$ for sources at $150$ and $200$ kpc.
For all distances, the \ac{CNN} maintains a \ac{FAP} no larger than $0.2\%$ for
both networks regarding magnetorotational waveforms. For neutrino-driven
signals, we show that H+L+VK has a \acp{TAP} of $39\%$ at $10$ kpc while it
is $6\%$ at $10$ kpc for HLVK with \acp{FAP} $\sim 0.1\%$. It should be noted that the results presented in this section are
averaged over all the data samples in the testing samples. The performance on
any individual waveform may vary depending on the morphology and the amplitudes
of the waveform.
